---
title: "Locality of various curves"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo = FALSE}
library(knitr)

knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center")

library(grid)
library(sfcurve)
```


For the Hilbert curve, calculate their Euclidean distance in the 2D space.

```{r}
p = hilbert_curve(level = 5)
library(ComplexHeatmap)
library(colorRamp2)

d_hc_2D = dist(p)
col_fun = colorRamp2(c(0, 20, 40), c("blue", "white", "red"))
ht1 = Heatmap(as.matrix(d_hc_2D), name = "dist", 
    col = col_fun, column_title = "Hilbert curve",
    cluster_rows = FALSE, cluster_columns = FALSE,
    show_row_names = FALSE, show_column_names = FALSE)
```

Similar for the H-curve, the Peano curve and the Meander curve.

```{r}
p = h_curve(iteration = 3)
d_h_2D = dist(p)
ht2 = Heatmap(as.matrix(d_h_2D), name = "dist", 
    col = col_fun, column_title = "H-curve",
    cluster_rows = FALSE, cluster_columns = FALSE,
    show_row_names = FALSE, show_column_names = FALSE)

p = peano_curve(level = 3)
d_pe_2D = dist(p)
ht3 = Heatmap(as.matrix(d_pe_2D), name = "dist", 
    col = col_fun, column_title = "Peano curve",
    cluster_rows = FALSE, cluster_columns = FALSE,
    show_row_names = FALSE, show_column_names = FALSE)

p = meander_curve(level = 3)
d_me_2D = dist(p)
ht4 = Heatmap(as.matrix(d_me_2D), name = "dist", 
    col = col_fun, column_title = "Meander curve",
    cluster_rows = FALSE, cluster_columns = FALSE,
    show_row_names = FALSE, show_column_names = FALSE)
```


```{r, fig.width = 8, fig.height = 8}
plot1 = grid.grabExpr(draw(ht1))
plot2 = grid.grabExpr(draw(ht2))
plot3 = grid.grabExpr(draw(ht3))
plot4 = grid.grabExpr(draw(ht4))
pushViewport(viewport(x = 0, y = 0.5, width = 0.5, height = 0.5, just = c("left", "bottom")))
grid.draw(plot1)
popViewport()
pushViewport(viewport(x = 0.5, y = 0.5, width = 0.5, height = 0.5, just = c("left", "bottom")))
grid.draw(plot2)
popViewport()
pushViewport(viewport(x = 0, y = 0, width = 0.5, height = 0.5, just = c("left", "bottom")))
grid.draw(plot3)
popViewport()
pushViewport(viewport(x = 0.5, y = 0, width = 0.5, height = 0.5, just = c("left", "bottom")))
grid.draw(plot4)
popViewport()
```

It seems all the four curves keep the locality very well that in the diagonals, the distance is 
quite small.

For a curve represented as a list of $n$ sequential points denoted as $(p_i)_n$, we calculate the
distance in the 2D space for points $p_i$ and $p_{i+k}$ where $k \ge 1$. We name $k$ as the gap
of two points.

$$
d_\mathrm{2D}(p_i, p_{i+k})
$$


```{r, fig.width = 8, fig.height = 4}
gap_dist = function(p, gap = 1) {
    n = nrow(p)
    loc1 = p[seq(1, n-gap), , drop = FALSE]
    loc2 = p[seq(1+gap, n), , drop = FALSE]
    sqrt( (loc1[, 1] - loc2[, 1])^2 + (loc1[, 2] - loc2[, 2])^2 )
}

par(mfrow = c(1, 2))
p = hilbert_curve(level = 6)
boxplot(lapply(1:20, function(i) gap_dist(p, i)), main = "Hilbert curve")
p = h_curve(iteration = 4)
boxplot(lapply(1:20, function(i) gap_dist(p, i)), main = "H-curve")
```

Similarly, we calculate such "gap distance" for the Peano curve and the Meander curve.
Because the numbers of points on the two curves are different from the Hilbert curve,
we rescale the size of the Peano curve and the Meander curve by their lengths.


```{r, fig.width = 8, fig.height = 4}
r = 9^4/4^6

par(mfrow = c(1, 2))
p = peano_curve(level = 4)
boxplot(lapply(1:20, function(i) gap_dist(p, i)/sqrt(r)), main = "Peano curve", names = round(1:20 /r, 1))
p = meander_curve(level = 4)
boxplot(lapply(1:20, function(i) gap_dist(p, i)/sqrt(r)), main = "Meander curve", names = round(1:20 /r, 1))
```

Last, we take the mean values from boxes and directly compare the corresponding gap distances.

```{r, fig.width = 5, fig.height = 5}
plot(NULL, xlim = c(1, 20), ylim = c(0, 6), xlab = "gaps", ylab = "mean gap dist")
p = hilbert_curve(level = 6)
lines(1:20, sapply(1:20, function(i) mean(gap_dist(p, i))), col = 2)
p = h_curve(iteration = 4)
lines(1:20, sapply(1:20, function(i) mean(gap_dist(p, i))), col = 7)
p = peano_curve(level = 4)

r = 9^4/4^6
lines(1:32 /r, sapply(1:32, function(i) mean(gap_dist(p, i)/sqrt(r))), col = 3)
p = meander_curve(level = 4)
lines(1:32 /r, sapply(1:32, function(i) mean(gap_dist(p, i)/sqrt(r))), col = 4)
legend("topleft", legend = c("hilbert", "h", "peano", "meander"), lty = 1, col = c(2, 7, 3, 4))
```

We can see, the Hilbert curve and the H-curve have the best locality. The 3x3 curve
performs worse than the 2x2 curves and the Peano curve performs the worst in the four 
curves.


