---
title: "3x3 Meander Curve"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo = FALSE}
library(knitr)

knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center")

library(grid)
library(sfcurve)
```


## Base patterns

The Meander curve has a different set of base units.

```{r, fig.width = 11.1, fig.height = 5.3, out.width = "100%"}
draw_rules_3x3_meander()
```

The expansions are very similar as the Hilbert curve, but it also has a
flipped version where the units can be flipped horizontally or vertically
on various levels.

```{r, fig.width = 11.1, fig.height = 5.3, out.width = "100%"}
draw_rules_3x3_meander(flip = TRUE)
```

On each level expansion, there are two transverse paths:

```{r, fig.width = 8, fig.height = 2.2}
plot_transverse_paths(SFC_RULES_3x3_MEANDER, SFC_RULES_3x3_MEANDER@rules$I[[1]])
```

## Curves

All possible forms of the Meander curve on level 3 taking `R(0)` as the seed:

```{r, fig.width = 8, fig.height = 4}
draw_multiple_curves(
    sfc_3x3_meander("R", code = "111"),
    sfc_3x3_meander("R", code = "112"),
    sfc_3x3_meander("R", code = "121"),
    sfc_3x3_meander("R", code = "122"),
    sfc_3x3_meander("R", code = "211"),
    sfc_3x3_meander("R", code = "212"),
    sfc_3x3_meander("R", code = "221"),
    sfc_3x3_meander("R", code = "222"),
    nrow = 2, lwd = 2)
```

The curves on level 3 taking the nine base patterns respectively:

```{r, fig.width = 9, fig.height = 9}
draw_multiple_curves(
    sfc_3x3_meander("I", code = "111"),
    sfc_3x3_meander("R", code = "111"),
    sfc_3x3_meander("L", code = "111"),
    sfc_3x3_meander("U", code = "111"),
    sfc_3x3_meander("B", code = "111"),
    sfc_3x3_meander("D", code = "111"),
    sfc_3x3_meander("P", code = "111"),
    sfc_3x3_meander("Q", code = "111"),
    sfc_3x3_meander("C", code = "111"),
    nrow = 3, lwd = 2)
```

## Unit flipping

Flipping (horizotal or vertical) can be applied by specifying the index of the hierarchical unit on the curve.

```{r, fig.width = 6, fig.height = 6}
p = sfc_3x3_meander("I", 11)
draw_multiple_curves(
    p,
    sfc_flip_unit(p, "1"),  # bottom left
    sfc_flip_unit(p, "2"),  # bottom middle
    sfc_flip_unit(p, c("6", "7", "8", "9")), # top left, four units
    nrow = 2
)
```

## Number of different forms

The total number of different forms of the Meander curve on level $k$ is composed by two parts:
$9 \cdot 4 \cdot 2^k$ is the number by applying the expanding rules, and $2^{\frac{9^k - 1}{8}}$
is the number of units on all levels where each unit can be flipped independently.


$$
9 \cdot 4 \cdot 2^k \cdot 2^{\frac{9^k - 1}{8}}
$$

## Seed as a sequence

```{r, fig.width = 3, fig.height = 3}
seed = sfc_sequence("IILILILI", rot = -90, universe = sfc_universe(SFC_RULES_3x3_MEANDER))
plot(seed, grid = TRUE)
```

```{r, fig.width = 6, fig.height = 6}
sfc_3x3_meander(seed, code = "111") |> plot(lwd = 2)
```
