---
title: "nxn-Curve"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo = FALSE}
library(knitr)

knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center")

library(grid)
library(sfcurve)
```

A curve with mode nxn can be generated by rules created for a specific type of nxn level-1 unit. Consider the following
4x4 level-1 unit which we call "the 4x4 Peano unit":

```{r, fig.width = 3, fig.height = 3, echo = FALSE}
sfc_unit("RIILLIIRRIILLIIR", rot = 90) |> sfc_grob() -> gb
gb$children[[1]]$gp$col = "black"
gb$children[[1]]$gp$lwd = 2
grid.draw(gb)
```

## 4x4 Peano curve

We will create rules for the expansion from level-0 to level-1 for the 4x4 Peano unit. We only cover the three major
base patterns of `I/R/L`. When creating the expansion rules, notice the following two requirements:

1: The in-direction and out-direction of the level-1 unit should be the same as these of the level-0 base patterns.
2: If the two level-1 units have the corner values of "12/21", "12" should always be the first one.

For each base pattern, there are two types of level-1 units. $I_1$ is:

```{r}
UNIVERSE_4x4_PEANO = c("I", "R", "L")
RULES_4x4_PEANO = list()
RULES_4x4_PEANO[["I"]][[1]] = sfc_unit("RIILLIIRRIILLIIR", rot = 0, universe = UNIVERSE_4x4_PEANO)
```

$I_2$ is basically a horizontal flip of $I_1$:

```{r}
RULES_4x4_PEANO[["I"]][[2]] = sfc_hflip(RULES_4x4_PEANO[["I"]][[1]])
```

$R_1$ is very similar as $R_2$ except the first letter and the last letter, also notice $R_2$ has a rotation of 270 because the in-direction
of the first base pattern should be bottom-up (90 degrees up).

```{r}
RULES_4x4_PEANO[["R"]][[1]] = sfc_unit("IIIRRIILLIIRRIIL", rot = 0, universe = UNIVERSE_4x4_PEANO)      
RULES_4x4_PEANO[["R"]][[2]] = sfc_rotate(sfc_unit("LIIRRIILLIIRRIII", 
                                 rot = 270, universe = UNIVERSE_4x4_PEANO), 90)
```

$L_1$ is a horizontal flip of $R_2$ and $L_2$ is a horizontal flip of $R_1$.

```{r}
RULES_4x4_PEANO[["L"]][[1]] = sfc_hflip(RULES_4x4_PEANO[["R"]][[2]])
RULES_4x4_PEANO[["L"]][[2]] = sfc_hflip(RULES_4x4_PEANO[["R"]][[1]])
```

Let's validate whether these level-1 units are in the correct forms and orientations:

```{r}
draw_multiple_curves(
	RULES_4x4_PEANO[["I"]][[1]],
	RULES_4x4_PEANO[["I"]][[2]],
	RULES_4x4_PEANO[["R"]][[1]],
	RULES_4x4_PEANO[["R"]][[2]],
	RULES_4x4_PEANO[["L"]][[1]],
	RULES_4x4_PEANO[["L"]][[2]],
	nrow = 2, extend = TRUE
)
```

Next we generate the `sfc_rules` object with the `sfc_rules()` function.

```{r}
SFC_RULES_4x4_PEANO = sfc_rules(rules = RULES_4x4_PEANO,
        name = "Peano 4x4",
        bases = BASE_LIST[UNIVERSE_4x4_PEANO])
SFC_RULES_4x4_PEANO
```

With the rules `SFC_RULES_4x4_PEANO`, we use the generator function `sfc_generator()` to generate two functions:

- `sfc_4x4_peano()`: main function for generating curves.
- `draw_rules_4x4_peano()`: draw the expansions from level-0 to level-1.

Here `"4x4_peano"` is the suffix of the two function names.

```{r}
sfc_generator(SFC_RULES_4x4_PEANO, "4x4_peano")
```

```{r, fig.width = 7.5, fig.height = 4}
draw_rules_4x4_peano()
```

```{r, fig.width = 8, fig.height = 4}
draw_multiple_curves(
	sfc_4x4_peano("I", "111"),
	sfc_4x4_peano("I", "121"),
	lwd = 2, nrow = 1
)
```


## 5x5 Meander curve

Similarly, we can generate the expansion rules for the 5x5 Meander curve.

```{r}
UNIVERSE_5x5_MEANDER = c("I", "R", "L")
RULES_5x5_MEANDER = list()
RULES_5x5_MEANDER[["I"]][[1]] = sfc_unit("IIIRIIRIRRLLILIILIIILIIIR", rot = 0, universe = UNIVERSE_5x5_MEANDER)
RULES_5x5_MEANDER[["I"]][[2]] = sfc_hflip(RULES_5x5_MEANDER[["I"]][[1]])
RULES_5x5_MEANDER[["L"]][[1]] = sfc_unit("IIIRIIRIRRLLILIILIIILIIII", rot = 0, universe = UNIVERSE_5x5_MEANDER)      
RULES_5x5_MEANDER[["L"]][[2]] = sfc_rotate(sfc_unit("LIIRIIRIRRLLILIILIIILIIIR", 
                                 rot = 270, universe = UNIVERSE_5x5_MEANDER), 90)
RULES_5x5_MEANDER[["R"]][[1]] = sfc_hflip(RULES_5x5_MEANDER[["L"]][[2]])
RULES_5x5_MEANDER[["R"]][[2]] = sfc_hflip(RULES_5x5_MEANDER[["L"]][[1]])

SFC_RULES_5x5_MEANDER = sfc_rules(rules = RULES_5x5_MEANDER,
        name = "Meander 5x5",
        bases = BASE_LIST[UNIVERSE_5x5_MEANDER])
```

And the curve generator:

```{r}
sfc_generator(SFC_RULES_5x5_MEANDER, "5x5_meander")
```

For simplicity, currently the flipped expansion rules are not supported in `sfc_generator()`.

```{r, fig.width = 9.5, fig.height = 5}
draw_rules_5x5_meander()
```

```{r, fig.width = 8, fig.height = 4}
draw_multiple_curves(
	sfc_5x5_meander("I", 11),
	sfc_5x5_meander("I", 21),
	lwd = 2, nrow = 1
)
```
