---
title: "Unit flipping"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo = FALSE}
library(knitr)

knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center")

library(grid)
library(sfcurve)
```

If the level-1 unit is not symmetric, there will be a flipped version of the expansion rules.
The flip can happen on the diagonal if the corner value of the unit is `11/22` or 
vertical/horizontal if the corner value of the unit is `12/21`.

We will use the Peano curve as an example because it is easier to identify different flipped
patterns by eyes.

```{r}
p = sfc_3x3_peano("I", rot = 90, level = 3)
```

## Flip individual units

Each unit on a certain level of the curve can be independently flipped. The unit is specified
by its "hierarchy index". The units are organized in a hierarchical way on the curve. See the following
examples:

```{r, fig.width = 9, fig.height = 3}
par(mfrow = c(1, 3))
sfcurve:::test_sfc_index(p, "") # the complete curve
sfcurve:::test_sfc_index(p, "4")
sfcurve:::test_sfc_index(p, "4:2")
```

`sfc_flip_unit()` flips a single unit or a list of units. Whether it is a diagonal flip or a 
vertical/horizontal flip is automatically detected.

```{r, fig.width = 9, fig.height = 3}
p2 = sfc_flip_unit(p, "7:3") # the most topright 3x3 unit 
p3 = sfc_flip_unit(p, as.character(1:9)) # every level-2 square units
draw_multiple_curves(
    p, p2, p3, 
    nrow = 1, lwd = 2)
```


## Flip level-1 units

There is a helper function `change_level1_unit_orientation()` which change the orientation
of all level-1 units simultaneously.

```{r, fig.width = 9, fig.height = 3}
p2 = change_level1_unit_orientation(p,  "vertical")
p3 = change_level1_unit_orientation(p,  "horizontal")
draw_multiple_curves(
    p, p2, p3, 
    nrow = 1, lwd = 2)
```

Note this type of flipping does not affect the higher level structure.

It internally applies `sfc_apply()` on the depth of `sfc_level(p)-1` (i.e. level = 1) which will introduced in later sections.

## Set flipping when creating the curve

As has already shown in the vignette ["Generalized Peano Curve"](peano_curve.html), `sfc_3x3_peano()` accepts
its argument `flip` as a self-defined function.

```{r, fig.width = 9, fig.height = 3}
p2 = sfc_3x3_peano("I", level = 3, flip = function(p) {
    p@rot %in% c(90, 270)
})
p3 = sfc_3x3_peano("I", level = 3, flip = function(p) {
    n = length(p)
    l = rep(FALSE, n)
    ind = seq(1, n/3)
    l[ind] = p@rot[ind] %in% c(90, 270)

    ind = seq(n/3+1, n/3*2)
    l[ind] = p@rot[ind] %in% c(0, 180)

    ind = seq(n/3*2+1, n)
    l[ind] = p@rot[ind] %in% c(90, 270)
    l
})
draw_multiple_curves(p, p2, p3, 
    nrow = 1, lwd = 2)
```

Being different from `change_level1_unit_orientation()` which is only applied on level 1, the self-defined
function `flip()` is appied to every level of the curve.

## Adjust flipping on a generated curve

We can also adjust the structure of the curve by `sfc_apply()` on a certain depth.
To better control the flipping patterns, we use `p2` where all level-1 units are
all vertical.

```{r, fig.width = 9, fig.height = 3}
p3 = sfc_apply(p2, 1, function(x, i) {
    if(i %% 2 == 1) {
        sfc_flip_unit(x)
    } else {
        x
    }
})
p4 = sfc_apply(p2, 2, function(x, i) {
    if(i %% 9 %in% 4:6) {
        sfc_flip_unit(x, to = "horizontal")
    } else {
        x
    }
})
draw_multiple_curves(
    p, p3, p4, 
    nrow = 1, lwd = 2)
```


## Seed as a sequence

`sfc_flip_unit()`, `change_level1_unit_orientation()` and `sfc_apply()` only work on curves with a single letter
as the seed because internally they use `sfc_index()` to get subunits and `sfc_index()` only works on square curves. 
Only the "flip" function works when the seed is a sequence.


```{r, fig.width = 6, fig.height = 3}
sfc_3x3_peano("IJ", level = 3, rot = 90, flip = function(p) {
    p@rot %in% c(90, 270)
}) |> plot()
```
